Do return series have power-law tails with the same index?

### 2017-06-02 papers/FX/draft.tex
Some revision on both papers.

\newcommand{\floor}[1]{
  \lfloor #1\rfloor
}
\newcommand{\ceil}[1]{
  \lceil #1\rceil
}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\allowdisplaybreaks
\begin{document}
\title{Do return series have power-law tails with the same index?}
\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
\date{\today}

@@

data from 4th January 2010 to 31 December 2014.

@@

are the estimated tail-indices from different series really distinct?

@@

Are there some {\em theoretical} reasons in favor of the fact that the tail-indices from different series are {\em not} 

@@

estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). {\red check} It turns out 
that the simultaneous estimation of the tail-index and the scale parameter are strongly related.

@@

is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991}. Gul derives the DA from first principles (axiomatic).

@@

\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
In this section, we assume the model \eqref{eq:1} for the tails of the marginal
\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
that this series constitutes a strictly stationary \seq . In what follows, we focus
on the left tail of the \ds , i.e., on the losses. 
However, it is common to present the tail-index estimators
for positive data. Therefore we will multiply the losses $X_t$ by
minus one, swapping the negative with the positive values.
For simplicity, we also suppress subscripts in the notation:
\beam\label{eq:1a}
\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
useful description of the risk. Alternatively, these parameters can be
used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
%With these motives in mind, we present a survey of the
%tail-indices and scale parameters of 3 sectors, namely ``Energy'',
%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
%index.

@@

order statistics. As  a matter of fact,

@@

i.e., $\hat \alpha$ is \asy ally unbiased and has variance

@@

verified based on the properties of the model. If they are not satisfied the 

@@

In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
estimates of the losses in the ``Energy'' and ``Consumer Staples''
sectors of the S\&P 500 index, as well as those of a 
large portion of losses in the ``Information Technology'' sector.
This fact indicates that the returns in each sector may 
have comparable tail-indices.
\par
Hoga's \cite{hoga:2016} test about the change of extreme quantiles
may provide some further insight about how similar these tail-indices are.
Different tail-indices are likely to result in different
extreme quantiles. Nevertheless, changes in the extreme quantiles may also
result from changing scale parameters in the tail. Therefore  we first investigate the scale
parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
\beao
  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
    0, K^{2\alpha}  \big)\,,
\end{eqnarray*}
where the tail-index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of $\hat K_k$
and $\hat K_k^{\hat\alpha}$ can be constructed.   
Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $\hat K_k$ (top) and $\hat K_k^{\hat
      \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
    of the S\&P 500 index. The points are the estimated values, the bars 
    the \asy\ 95\%-confidence intervals; the confidence bands of the
    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
    shown in Figure~\ref{fig:1}. 
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd row of
Figure~\ref{fig:sectors_parameters}.
\par
Secondly, it appears that there is positive dependence between the  values of $\alpha$ and $K$. 
This prompts Pareto tails of the return
distributions. For more explanations, see
Section~\ref{sec:pareto_tail}.
\par
Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
average, the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are  larger
than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
a larger value of $K$ implies that large losses are more
probable. Thus one can conclude that these two sectors are
considerably riskier than the ``Consumer Staples'' sector. {\red Let
  Casper decide about this sentence.} This is of course a confirmation
of one's economic instinct.

Yet another indication from Figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the dependence  between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, moving along
a curve of equal preference in the direction of increasing $\alpha$, $K$ also
increases. So the strong positive dependence seen in the ``Energy''
sector suggests these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and do not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

@@

    sectors of S\&P 500. The test statistic in \eqref{eq:x1} is based

@@

    The same number of upper order statistics are used for both tests.}

@@

It indicates that tail-indices of equities in the

@@

extreme quantiles change as well. We will test for a change of tail
index or scale parameter in this indirect way.

@@

% \begin{figure}[htb!]
%   \begin{minipage}{0.5\linewidth}
%     \includegraphics[
%       width=\textwidth,
%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
%     ]{Hoga_Energy_pair.pdf}
%   \end{minipage}\hfill
%   \begin{minipage}{0.5\linewidth}
%     \includegraphics[
%       width=\textwidth,
%       trim={0.3cm, 1cm, 1cm, 1cm}, clip
%     ]{Hoga_CS_pair.pdf}
%   \end{minipage}
%   \caption{{\red check the symbols on the axes. Some are too long.} Test for changing tail-index or scale parameter based on concatenated series in the ``Energy'' and
%     ``Consumer Staples'' sectors of
%     S\&P 500. The green, blue and red colors represent, respectively, the cases when  
%     the 85\%, 90\% and 95\% quantiles of the limit distribution of $T_n$ is exceeded.
%     Grey points stand for pairs for which $T_n$ is below
%     the 85\% \asy\ quantile;
%     black points represent pairs for which
%     the computation of $T_n$ fails for given precision
%     requirements and time limits.}
%   \label{fig:Hoga_pair}
% \end{figure}

To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size
$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
distinct degrees of freedom. The results are shown in
Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
larger the minimum tail-index in the concatenated pair.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \end{minipage}
  \caption{
    {\em Left}: Test  of concatenated $t$-samples with different
    degrees of freedom ($\alpha$). Numbers on the axes are the degrees of freedom
    in the subsamples. For an interpretation of the colored bullets,
    see the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
    of the test. In particular, if both degrees of freedom are
    relatively large it loses the capability of distinguishing between the \ds s.
    {\em Right}: Comparison of the \asy\ distribution of the test statistic
    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
    of freedom.
  }
  \label{fig:t_sim_pair}
\end{figure}

A major problem in this test is the \asy\ distribution of the test statistic under the
null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
To find out about this problem we compared the \ds s of $T_n$ 
with $n=1304$ for $t$-distributed $X_t$ with  3 and 4 degrees of freedom ($\alpha$) and
the limit \ds\ of $T_n$.
The estimated density functions are shown on the right of Figure~\ref{fig:t_sim_pair}.
As seen in the graph, the asymptotic distribution assigns
significantly more mass to the tail than the \ds s of $T_n$ do. For
comparison, we list a few quantiles of these \ds s in
Table~\ref{tab:HogaAsymptotic}, showing major differences between the
\asy\ and finite-sample \ds s.
\begin{table}[htb!]
  \centering
  \begin{tabular}{l|c|c|c|r}
    & \multicolumn{4}{c}{Quantiles} \\[2mm]
    \hline
    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
    \hline
    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
  \end{tabular}
  \caption{Quantiles of the test statistic $T_n$ for $n=1304$
    $t$-distributed samples with 3 and 4 degrees of freedom ($\alpha$)
    as well as the corresponding quantiles for the limiting \ds\ of
    $T_n$. In particular, there are huge differences between the three
    \ds s  for the higher quantiles. 
    }
  \label{tab:HogaAsymptotic}
\end{table}

We also acknowledge that the test is designed to detect a sudden big
change and cannot detect changes that are small but accumulate over
time. This is due to the fact that $T_n$ is not invariant under
permutation of the observations ($X_t$).
If the observations are iid, the null hypothesis is of 
course not rejected with or without permutation; If instead the
observations are not iid but rather have small changes of the
tail-index over time, the null hypothesis may be rejected with
permutation but not rejected without. The latter may happen when the
changes in the tail-index are small but just big enough for the
accumulated change between two time points to exceed the
critical value. In other words, the test cannot detect slow changes.

We illustrate this point in figure \ref{fig:PairTest:permuted} where
we apply a random permutation of observations to each series before
concatenating them and running the test.
\begin{figure}[htb!]
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_Energy_pair_permuted.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_CS_pair_permuted.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_IT_pair_permuted.pdf}
  \end{minipage}
  \caption{
    Test for changing tail-index or scale parameter
    of losses
    using Hoga's test based on concatenated series of pairs of
    stocks.
    A random permutation is applied to the observations of each
    series.
    The green, blue and red points
    correspond to pairs of stock in a sector 
    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
    $95\%$-quantile of the limit \ds .
    Grey points stand for pairs for which the test statistic is below
    the \asy\ $85\%$-quantile. Black points represent 
    pairs for which the computation of $T_n$ fails for given precision  
    requirements and time limits.
  }
  \label{fig:PairTest:permuted} 
\end{figure}

\newpage
\section{Some theoretical arguments for equality of tail-indices}\setcounter{equation}{0}\label{sec:2}
\subsection{Multivariate GARCH models whose components have  equal tail-indices}\label{subsec:garch}
Among the models for returns the generalized autoregressive conditionally heteroscedastic (GARCH) model
is certainly most popular because it is parsimonious, captures various of the stylized facts of real return data
and can also be modified in various directions to capture specific behavior of \ts\ such a asymmetry, skewness, long memory; 
see for example Andersen et al. \cite{andersen:davis:kreiss:mikosch:2009}, Part 1, for a collection of results
on GARCH-type models.
The original  {\em univariate} GARCH model of Bollerslev \cite{bollerslev:1986} is a \sv\ model of the type $X_t=\sigma_t\,Z_t$, where 
$(Z_t)$ is an iid mean-zero unit-variance \seq . In the simple case of a \garch\ the {\em squared volatility} satisfies the {\em \sre} 
\beam\label{eq:5}
\sigma_t^2= \alpha_0+\alpha_1 X_{t-1}^2+\beta_1\sigma_{t-1}^2=\alpha_0+(\alpha_1Z_{t-1}^2+\beta_1)\sigma_{t-1}^2\,,\quad t\in\bbz\,.
\eeam
Here $\alpha_0>0$, $\alpha_1,\beta_1$ are non-negative constants. For suitable choices of $\alpha_1,\beta_1$ the equation \eqref{eq:5}
can be solved and the solution $(\sigma_t^2)$ constitutes a strictly stationary \seq , implying that $(X_t)$ 
is strictly stationary itself. A remarkable property of the process $(\sigma_t)$ is that it has a power-law tail 
of the form
\beam\label{eq:6}
\P(\sigma_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
for some positive $c>0$ and a positive tail-index $\alpha$ which is the unique solution of the equation 
$\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
provided that the solution exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
Buraczewski et al. \cite{buraczewski:damek:mikosch:2016}. The latter result ensures power-law tails  for
the strictly stationary  solution $(Y_t)$ to the \sre\ 
\beam\label{eq:7}
Y_t= A_t\,Y_{t-1}+B_t\,,\qquad t\in\bbz\,,
\eeam
for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^{\alpha/2}]=1$.  
In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2+\beta_1$ and $B_t=\alpha_0$ to achieve
\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016}, p.~275)
it follows that
\beao
\P(\pm X_t>x)\sim \E[(Z_t)_\pm^\alpha]\,\P(\sigma_t>x)\,,\qquad \xto\,, 
\eeao
implying power-laws for the right and left tails of $X_t$ caused by the power-law tail of $\sigma_t$.

@@

\right)\,.

@@

@@

are degenerate since they do not allow for a linear relationship

@@

$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$ and
$\alpha, \beta, K, K'$ are independent of each other. We shall also
use the density function of $X$, which we define as
$f_X(x) = {\pd F_X(x) \over \pd x}$.
Then we have

@@

We observe the following property which also applies to \eqref{eq:xxie1.0}. 

@@

 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.

@@

The proof is given in Appendix~\ref{sec:thrmI_proof}. 

@@

>>>>

\newcommand{\floor}[1]{
  \lfloor #1\rfloor
}
\newcommand{\ceil}[1]{
  \lceil #1\rceil
}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\allowdisplaybreaks
\begin{document}
\title{Do return series have power-law tails with the same index?}
t\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
\date{\today}

@@

data from 4 January 2010 to 31 December 2014.

@@

are the tail-indices from different series really distinct?

@@

Are there some {\em theoretical} reasons supporting the fact that the tail-indices from different series are {\em not} 

@@

estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). 
%{\red check} It turns out 
%that the simultaneous estimation of the tail-index and the scale parameter are strongly related.

@@

is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991} who derived DA from first principles (axiomatic).

@@

\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
In this section, we assume the model \eqref{eq:1} for the tails of the marginal
\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
that this series constitutes a strictly stationary \seq . In what follows, we focus
on the left tail of the \ds , i.e., on the losses. 
However, it is common to present the tail-index estimators
for positive data. Therefore we will multiply the losses $X_t$ by
minus one, swapping the negative with the positive values.
For simplicity, we also suppress subscripts in the notation:
\beam\label{eq:1a}
\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
useful description of the risk, for example in terms of high quantiles such as
Value-at-Risk. Alternatively, these parameters can be
used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
%With these motives in mind, we present a survey of the
%tail-indices and scale parameters of 3 sectors, namely ``Energy'',
%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
%index.

@@

upper order statistics. As  a matter of fact,

@@

i.e., $\hat \alpha_k$ is \asy ally unbiased and has variance

@@

verified based on the theoretical properties of the model. If they are not satisfied the 

@@

In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
estimates of the losses in the ``Energy'' and ``Consumer Staples''
sectors of the S\&P 500 index, as well as those of a 
large portion of losses in the ``Information Technology'' sector.
This fact indicates that the returns in each sector may 
have comparable tail-indices.
\par
Hoga's \cite{hoga:2016} test about the change of extreme quantiles
in a sample
may provide some further insight about how similar these tail-indices are.
Different tail-indices are likely to result in different
extreme quantiles. Nevertheless, changes in the extreme quantiles may also
result from changing scale parameters in the tail. Therefore  we first investigate the scale
parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
\beao
  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
    0, K^{2\alpha}  \big)\,,
\end{eqnarray*}
where the tail-index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of $\hat K_k$
and $\hat K_k^{\hat\alpha}$ can be constructed.   
Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $\hat K_k$ (top) and $\hat K_k^{\hat
      \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
    of the S\&P 500 index. The estimates are ordered according to the 
    corresponding estimated $\alpha$-values.
    The points are the estimated values, the bars 
    the \asy\ 95\%-confidence intervals; the confidence bands of the
    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
    shown in Figure~\ref{fig:1}. 
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd row of
Figure~\ref{fig:sectors_parameters}.
\par
Secondly, it appears that there is positive dependence between the
values of $\alpha$ and $K$.
As argued in Section~\ref{sec:pareto_tail}, this is consistent with
the assumption that the return series have Pareto tails on both sides
with the tail parameters on each side independent of those on the
other.

\par
Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
average, the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are  larger
than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
a larger value of $K$ implies that large losses are more
probable. Thus one can conclude that these two sectors are
considerably riskier than the ``Consumer Staples'' sector. {\red Casper:
we leave this sentence to your discretion.} This is of course a
confirmation of one's economic instinct.

Yet another indication from Figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the dependence  between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, when moving along
a curve of equal preference in the direction of increasing $\alpha$, the parameter $K$ also
increases. So the strong positive dependence seen in the ``Energy''
sector suggests that these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and do not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

@@

    sectors of S\&P 500. The test statistic in $\hat\alpha_X-\hat \alpha_Y$ is based

@@

    The same number (50) of upper order statistics is used for both tests.}

@@

They indicate that tail-indices of equities in the

@@

extreme quantiles change as well. We will test for a change of tail-index or scale parameter in this indirect way.

@@

To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size
$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
distinct degrees of freedom. The results are shown in
Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
larger the minimum tail-index in the concatenated pair.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \end{minipage}
  \caption{
    {\em Left}: Test  of concatenated $t$-samples with different
    degrees of freedom $\alpha$. Numbers on the axes are the degrees of freedom
    in the subsamples. For an interpretation of the colored bullets,
    see the caption for the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
    of the test. In particular, if both degrees of freedom are
    relatively large it loses the capability of distinguishing between the \ds s.
    {\em Right}: Comparison of the \asy\ distribution of the test statistic
    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
    of freedom.
  }
  \label{fig:t_sim_pair}
\end{figure}

A major problem of this test is the \asy\ distribution of the test statistic under the
null hypothesis.  The rate at which the finite-sample distribution tends to its limit is not known.
To find out about this problem we compared the \ds s of $T_{1304}$ 
for $t$-distributed $X_t$ with  $\alpha=3$ and $\alpha=4$ degrees of freedom with
the limit \ds\ of $T_n$.
The estimated density functions are shown on the right of Figure~\ref{fig:t_sim_pair}.
As seen in the graph, the asymptotic distribution assigns
significantly more mass to the tail than the \ds s of $T_n$ do. For
comparison, we list a few quantiles of these \ds s in
Table~\ref{tab:HogaAsymptotic}, showing major differences between the
\asy\ and finite-sample \ds s.
\begin{table}[htb!]
  \centering
  \begin{tabular}{l|c|c|c|r}
    & \multicolumn{4}{c}{Quantiles} \\[2mm]
    \hline
    Distribution of $X_t$& 80\% & 85\% & 90\% & 95\% \\
    \hline
    Asymptotic & 47.48 & 59.61 & 78.90 & 113.12 \\
    t(3)  & 23.30 & 28.27 & 35.04 & 46.75 \\
    t(4)  & 25.76 & 31.13 & 38.81 & 56.32\\[2mm]
  \end{tabular}
  \caption{Quantiles of the test statistic $T_n$ for $n=1304$
    $t$-distributed samples with $\alpha=3$ and $\alpha=4$ 
degrees of freedom
    as well as the corresponding quantiles for the limiting \ds\ of
    $T_n$. In particular, there are huge differences between the three
    \ds s  for the higher quantiles. 
    }
  \label{tab:HogaAsymptotic}
\end{table}

Figure~\ref{fig:PairTest:permuted} points at another shortcoming
of the test: we show $T_n$ for an 
arbitrarily chosen random permutation of the concatenated data from two 
different stocks. In this case the null hypothesis that two series
have the same tail-index is rejected much more often, as a comparison
with the bottom graphs of
Figure~\ref{fig:PairTest} shows. If the data in the concatenated 
series were iid a random permutation would not change the 
\ds\ of $T_n$. Thus the value of the test statistic $T_n$  
strongly depends on the dependence structure of the underlying data
and therefore a test based on $T_n$ may be misleading.

\begin{figure}[htb!]
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_Energy_pair_permuted.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_CS_pair_permuted.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[
      width=\textwidth,
      trim={0.3cm, 0.8cm, 1cm, 0.6cm}, clip
    ]{Hoga_IT_pair_permuted.pdf}
  \end{minipage}
  \caption{
    Test for changing tail-index or scale parameter
    of losses
    using Hoga's test based on concatenated series of pairs of
    stocks.
    A random permutation is applied to the observations of each
    series.
    The green, blue and red points
    correspond to pairs of stock in a sector 
    when the test statistic $T_n$ exceeds the $85\%$-, $90\%$-,
    $95\%$-quantile of the limit \ds .
    Grey points stand for pairs for which the test statistic is below
    the \asy\ $85\%$-quantile. Black points represent 
    pairs for which the computation of $T_n$ fails for given precision  
    requirements and time limits.
  }
  \label{fig:PairTest:permuted} 
\end{figure}

\newpage
\section{Some theoretical arguments for equality of tail-indices}\setcounter{equation}{0}\label{sec:2}
\subsection{Multivariate GARCH models whose components have  equal tail-indices}\label{subsec:garch}
Among the models for returns the generalized autoregressive conditionally heteroscedastic (GARCH) model
is certainly most popular because it is parsimonious, captures various of the stylized facts of real return data
and can also be modified in various directions to capture specific behavior of \ts\ such a asymmetry, skewness, long memory; 
see for example Andersen et al. \cite{andersen:davis:kreiss:mikosch:2009}, Part 1, for a collection of results
on GARCH-type models.
The original  {\em univariate} GARCH model of Bollerslev \cite{bollerslev:1986} is a \sv\ model of the type $X_t=\sigma_t\,Z_t$, where 
$(Z_t)$ is an iid mean-zero unit-variance \seq . In the simple case of a \garch\ the {\em squared volatility} satisfies the {\em \sre} 
\beam\label{eq:5}
\sigma_t^2= \alpha_0+\alpha_1 X_{t-1}^2+\beta_1\sigma_{t-1}^2=\alpha_0+(\alpha_1Z_{t-1}^2+\beta_1)\sigma_{t-1}^2\,,\quad t\in\bbz\,.
\eeam
Here $\alpha_0>0$, $\alpha_1,\beta_1$ are non-negative constants. For suitable choices of $\alpha_1,\beta_1$ the equation \eqref{eq:5}
can be solved and the solution $(\sigma_t^2)$ constitutes a strictly stationary \seq , implying that $(X_t)$ 
is strictly stationary itself. A remarkable property of the process $(\sigma_t)$ is that it has a power-law tail 
of the form
\beam\label{eq:6}
\P(\sigma_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
for some positive $c>0$ and a positive tail-index $\alpha$ which is the unique solution of the equation 
$\E [(\alpha_1 Z_1^2+\beta_1)^{\alpha/2}]=1$
provided that the solution exists and some mild assumptions on the \ds\ of $Z_t$ hold. This result
follows by an application of the Kesten-Goldie theorem; see Kesten \cite{kesten:1973}, Goldie \cite{goldie:1991}, cf. 
Buraczewski et al. \cite{buraczewski:damek:mikosch:2016} for a 
recent textbook treatment. The latter result ensures power-law tails  for
the strictly stationary  solution $(Y_t)$ to the \sre\ 
\beam\label{eq:7}
Y_t= A_t\,Y_{t-1}+B_t\,,\qquad t\in\bbz\,,
\eeam
for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^{\alpha/2}]=1$.  
In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2+\beta_1$ and $B_t=\alpha_0$ to achieve
\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016}, p.~275)
it follows that
\beao
\P(\pm X_t>x)\sim \E[(Z_t)_\pm^\alpha]\,\P(\sigma_t>x)\,,\qquad \xto\,, 
\eeao
implying power-laws for the right and left tails of $X_t$ caused by the power-law tail of $\sigma_t$.

@@

\right)\,,

@@

for positive $\alpha_{0i}$ and suitable non-negative 
$\alpha_{ij},\beta_{ij}$, $i,j=1,2$.

@@

are not natural: they are degenerate since they do not allow 
for a linear relationship

@@

$\alpha, \beta > 0$, $K,K' > 0$, $0 < p < 1$. 
We also write $f_X$ for the density function of $F_X$.
We have

@@

We observe the following property 
whose proof is given in Appendix~\ref{sec:thrmI_proof}. 
% which also applies to the special case 
%\eqref{eq:xxie1.0}. 

@@

Assume the two-sided Pareto model \eqref{eq:pareto},
that there is no \fct al relationship between $\alpha,K$ and $\beta,K'$
and the utility \fct\ $u$ is increasing and differentiable. Then
$\frac{\pd \wt u_{\rm max}}{\pd \alpha} > 0$ and
$\frac{\pd \wt u_{\rm max}}{\pd K} < 0$.

@@

@@

U_{\text{all}}(x) =
